---
title: "Mapping U.S. presidential elections results by state"
author: "by BPG"
date: "2/11/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(data.table)
library(tidyr)
library(sp)
library(sf)
library(stringr)
library(spData)
library(tmap)
library(elections)

```


## How to start using R Markdown

You need to install[pandoc](pandoc.org) and two packages:

```{r packages needed, eval = FALSE}
install.packages("rmarkdown")
install.packages("knitr")
```


## Loading needed data

```{r data loading}

presidentsdata = fread('/Users/Belka/Downloads/1976-2020-president.csv', dec=",") 
glimpse(presidentsdata)

data(eldat)
glimpse(eldat)

#Spatial Data
glimpse(us_states)
```


## Data

- Results of all presidential elections from `r min(presidentsdata$year)` to `r max(presidentsdata$year)` by state level
- Data table with elected presidents since `r min(eldat$electionyear)`
- Spatial data of state boundaries 


## Data transformation

At this stage we transform data to make it applicable for our purposes

```{r data processing}
presidentsdata_cleaned = presidentsdata %>% select(year,state,candidate, party_simplified, candidatevotes)  %>% group_by(year, state)  %>% top_n(1, candidatevotes) # selecting needed information

head(presidentsdata_cleaned)

years = unique(presidentsdata_cleaned$year) 

presidentsdata_cleaned = presidentsdata_cleaned %>% mutate(party_cand = paste0(party_simplified, ' (', candidate ,')'))  %>% select(year, state, party_cand) 

head(presidentsdata_cleaned)

presidentsdata_cleaned_wide = presidentsdata_cleaned %>% pivot_wider(names_from = year, values_from = party_cand) %>% na.omit() %>% mutate(state = str_to_title(state)) 

head(presidentsdata_cleaned_wide)

presidentsdatasf = us_states %>% left_join(presidentsdata_cleaned_wide, by = c("NAME" = "state"))

head(presidentsdatasf)

#Second table
presidentwin = eldat %>% filter(electionyear %in% years) %>% select(electionyear, winner,winnerparty)

pr2020 = c('2020', 'Joe Biden', 'Dem.' ) # creating results for 2020 elections
presidentwin  = presidentwin %>% rbind(pr2020) #  adding 2020 to the table

head(presidentwin)

```


## Creating maps for each year

```{r creating set of maps}

map = list()

for (i in 1:12) {  
  map[[i]] = tm_shape(presidentsdatasf) +
    tm_polygons(colnames(presidentsdatasf)[i+6],
                palette = c('royalblue2', 'brown1'),
                border.col = "white",
                lwd = 0.3, 
                showNA = FALSE)+
    tm_legend(legend.position = c("left", "bottom"))+
    tm_layout(inner.margins = c(0.1, 0.1, 0.1, 0.1), fontfamily = 'Muller', frame = FALSE)}


```


# Examples of maps

```{r examples of maps}

map[[1]]
map[[12]]

```


